Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2024-08-11 Thesis/03-chap3.qmd
update model violations index section

In this chapter, we develop computer vision models and integrate them into the residual plots diagnostics workflow, addressing the need for an automated visual inference. The chapter is structured as follows. @sec-model-specifications discusses various specifications of the computer vision models. @sec-model-distance-between-residual-plots defines the distance measure used to measure model violations. @sec-model-distance-estimation explains how the computer vision models estimate this distance measure. @sec-model-statistical-testing covers the statistical testing based on the estimated distance. Sections [-@sec-model-data-generation], [-@sec-model-architecture], and [-@sec-model-training] detail the data preparation, model architecture, and training process, respectively. The results are presented in @sec-model-results. Finally, we conclude with a discussion of our findings and propose ideas for future research directions.

@@

## Model Violations Index {#sec:model-violations-index}

$\hat{D}$ is an estimator of the difference between the true residual distribution and the reference residual distribution. This difference primarily arises from deviations in model assumptions. The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violations index (MVI).

@@

::: {#tbl-mvi .cell tbl-cap='Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.'}
::: {.cell-output-display}

@@

```{r}
#| tbl-cap: "Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10."
#| label: tbl-mvi
data.frame(degree = c("Strong", "Moderate", "Weak"),
           range = c("$\\text{MVI} > 8$", "$6 < \\text{MVI} < 8$", "$\\text{MVI} < 6$")) %>%
  kableExtra::kable(booktabs = TRUE,
                    escape = FALSE,
                    col.names = c("Degree of model violations", "Range ($C$ = 10)"),
                    align = "lc") %>%
  kableExtra::kable_styling(latex_options = c("hold_position"))
```

@@

```{r}
#| label: tbl-hyperparameter
#| tbl-cap: "Name of hyperparameters and their correspoding domain for the computer vision model."
data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[$10^{-8}$, $10^{-1}$]")) %>%
  kableExtra::kable(booktabs = TRUE)
```

@@

## Examples

In this section, we present the performance of trained computer vision model on three example datasets. These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in @fig-false-finding, along with the Boston housing dataset [@harrison1978hedonic], and the "dino" datasets from the `datasaurus` R package [@datasaurus].

@@

```{r fig-false-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots. (A) The true residual plot exhibiting a "left-triangle" shape. (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.  (C) The density plot of estimated distnace for null plots and bootstrapped plots. The green area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots. The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$. (D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model.  ', fig.height = 8}
p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
p2 <- readRDS(here::here("cached_data/fig1_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(HETER_MODEL$test(dat)$p_value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(observed~hat(D)))) +
  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
pca_result <- my_vi$feature_pca()
pc1_per <- (attributes(pca_result)$sdev^2/256)[1]
pc2_per <- (attributes(pca_result)$sdev^2/256)[2]
p4 <- pca_result %>%
  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
  mutate(set = ifelse(set == "null", "Null", set)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = 10) +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  xlab(glue::glue("PC1 ({scales::percent(pc1_per)})")) +
  ylab(glue::glue("PC2 ({scales::percent(pc2_per)})")) +
  labs(col = "") +
  theme(legend.position = "bottom") +
  ggtitle("(D) PCA for extracted features") +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))

@@

```{r fig-boston-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Boston housing fitted model evaluted on 200 null plots and 200 bootstrapped plots. (A) The true residual plot exhibiting a "U" shape. (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.  (C) The density plot of estimated distnace for null plots and bootstrapped plots. The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots. The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$. (D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model. ', fig.height = 8}
p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
p2 <- readRDS(here::here("cached_data/boston_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(observed~hat(D)))) +
  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
pca_result <- my_vi$feature_pca()
pc1_per <- (attributes(pca_result)$sdev^2/256)[1]
pc2_per <- (attributes(pca_result)$sdev^2/256)[2]
p4 <- pca_result %>%
  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
  mutate(set = ifelse(set == "null", "Null", set)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = 7) +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  xlab(glue::glue("PC1 ({scales::percent(pc1_per)})")) +
  ylab(glue::glue("PC2 ({scales::percent(pc2_per)})")) +
  labs(col = "") +
  theme(legend.position = "bottom") +
  ggtitle("(D) PCA for extracted features") +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))

@@

### Datasaurus

@@

The "dino" dataset within the `datasaurus` R package exemplifies this scenario. With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur", as displayed in [@fig-dino-check]A. Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in @fig-dino-lineup. A visual test conducted by humans would undoubtedly reject $H_0$.

@@

```{r fig-dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Datasaurus fitted model evaluted on 200 null plots and 200 bootstrapped plots. (A) The residual plot exhibits a "dinosaur" shape. (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.  (C) The density plot of estimated distnace for null plots and bootstrapped plots. The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots. The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$. (D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model.', fig.height = 8}

p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot")
p2 <- readRDS(here::here("cached_data/dino_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("RESET test p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}\n Breusch-Pagan test p-value = {format(lmtest::bptest(mod)$p.value, digits = 3)} \n Shapiro-Wilk test p-value = {format(shapiro.test(mod$residuals)$p.value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(observed~hat(D)))) +
  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
pca_result <- my_vi$feature_pca()
pc1_per <- (attributes(pca_result)$sdev^2/256)[1]
pc2_per <- (attributes(pca_result)$sdev^2/256)[2]
p4 <- pca_result %>%
  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
  mutate(set = ifelse(set == "null", "Null", set)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = -10) +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  xlab(glue::glue("PC1 ({scales::percent(pc1_per)})")) +
  ylab(glue::glue("PC2 ({scales::percent(pc2_per)})")) +
  labs(col = "", shape = "") +
  theme(legend.position = "bottom", legend.box="vertical") +
  ggtitle("(D) PCA for extracted features") +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))

@@

In this chapter, we have introduced a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference residual distribution assumed under correct model specification. This distance measure effectively captures the magnitude of model violations in misspecified models. We propose a computer vision model to estimate this distance, utilizing the residual plot of the fitted model as input. The resulting estimated distance serves as the foundation for constructing a single Model Violation Index (MVI), facilitating the quantification of various model violations.

>>>>

In this chapter, we develop computer vision models and integrate them into the residual plots diagnostics workflow, addressing the need for an automated visual inference. The chapter is structured as follows. @sec-model-specifications discusses various specifications of the computer vision models. @sec-model-distance-between-residual-plots defines the distance measure used to detect model violations, while @sec-model-distance-estimation explains how the computer vision models estimate this distance measure. @sec-model-statistical-testing covers the statistical tests based on the estimated distance, and @sec-model-violations-index introduces a Model Violations Index, which offers a quicker and more convenient assessment. Sections [-@sec-model-data-generation], [-@sec-model-architecture], and [-@sec-model-training] detail the data preparation, model architecture, and training process, respectively. The results are presented in @sec-model-results. Example dataset applications are discussed in @sec-examples. Finally, we conclude with a discussion of our findings and propose ideas for future research directions.

@@

## Model Violations Index {#sec-model-violations-index}

While statistical testing is a powerful tool for detecting model violations, it can become cumbersome and time-consuming when quick decisions are needed, particularly due to the need to evaluate numerous null plots. In practice, a more convenient and immediate method for assessing model performance is often required. This is where an index, such as the Model Violations Index (MVI), becomes valuable. It offers a straightforward way to quantify deviations from model assumptions, enabling rapid assessment and easier comparison across models.

The estimator $\hat{D}$ measures the difference between the true residual distribution and the reference residual distribution, a difference primarily arises from deviations in model assumptions. The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violations index (MVI).

@@

::: {#tbl-mvi .cell tbl-cap='Degree of model violations or the strength of the visual signals according to the Model Violations Index (MVI). The constant $C$ is set to be 10.'}
::: {.cell-output-display}

@@

```{r}
#| tbl-cap: "Degree of model violations or the strength of the visual signals according to the Model Violations Index (MVI). The constant $C$ is set to be 10."
#| label: tbl-mvi
data.frame(degree = c("Strong", "Moderate", "Weak"),
           range = c("$\\text{MVI} > 8$", "$6 < \\text{MVI} < 8$", "$\\text{MVI} < 6$")) %>%
  kableExtra::kable(booktabs = TRUE,
                    escape = FALSE,
                    col.names = c("Degree of model violations", "Range ($C$ = 10)"),
                    align = "lc")
```

@@

::: {#tbl-hyperparameter .cell tbl-cap='Name of hyperparameters and their correspoding domain for the computer vision model.'}
::: {.cell-output-display}

`````{=html}
<table>
 <thead>
  <tr>
   <th style="text-align:left;"> Hyperparameter </th>
   <th style="text-align:left;"> Domain </th>
  </tr>
 </thead>
<tbody>
  <tr>
   <td style="text-align:left;"> Number of base filters </td>
   <td style="text-align:left;"> {4, 8, 16, 32, 64} </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Dropout rate for convolutional blocks </td>
   <td style="text-align:left;"> [0.1, 0.6] </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Batch normalization for convolutional blocks </td>
   <td style="text-align:left;"> {false, true} </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Type of global pooling </td>
   <td style="text-align:left;"> {max, average} </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Ignore additional inputs </td>
   <td style="text-align:left;"> {false, true} </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Number of units for the fully-connected layer </td>
   <td style="text-align:left;"> {128, 256, 512, 1024, 2048} </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Batch normalization for the fully-connected layer </td>
   <td style="text-align:left;"> {false, true} </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Dropout rate for the fully-connected layer </td>
   <td style="text-align:left;"> [0.1, 0.6] </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Learning rate </td>
   <td style="text-align:left;"><span data-qmd = "[$10^{-8}$, $10^{-1}$]"></span> </td>
  </tr>
</tbody>
</table>

```{r}
#| label: tbl-hyperparameter
#| tbl-cap: "Name of hyperparameters and their correspoding domain for the computer vision model."
data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[$10^{-8}$, $10^{-1}$]")) %>%
  kableExtra::kable(booktabs = TRUE, escape = FALSE)
```

@@

## Examples {#sec-examples}

In this section, we present the performance of trained computer vision model on three example datasets. These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in @fig-false-finding, along with the Boston housing dataset [@harrison1978hedonic], and the "dino" datasets from the `datasauRus` R package [@datasaurus].

@@

```{r fig-false-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots. (A) The true residual plot exhibiting a "left-triangle" shape. (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.  (C) The density plot of estimated distnace for null plots and bootstrapped plots. The green area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots. The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$. (D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model.  ', fig.height = 8}
p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle(glue::glue("(A) Residual plot (MVI = {format(my_vi$check_result$observed$vss - log(nrow(my_vi$get_fitted_and_resid())) + 10, digits = 3)}, weak)")) 
p2 <- readRDS(here::here("cached_data/fig1_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(HETER_MODEL$test(dat)$p_value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(observed~hat(D)))) +
  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
pca_result <- my_vi$feature_pca()
pc1_per <- (attributes(pca_result)$sdev^2/256)[1]
pc2_per <- (attributes(pca_result)$sdev^2/256)[2]
p4 <- pca_result %>%
  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
  mutate(set = ifelse(set == "null", "Null", set)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = 10) +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  xlab(glue::glue("PC1 ({scales::percent(pc1_per)})")) +
  ylab(glue::glue("PC2 ({scales::percent(pc2_per)})")) +
  labs(col = "") +
  theme(legend.position = "bottom") +
  ggtitle("(D) PCA for extracted features") +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))

@@

```{r fig-boston-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Boston housing fitted model evaluted on 200 null plots and 200 bootstrapped plots. (A) The true residual plot exhibiting a "U" shape. (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.  (C) The density plot of estimated distnace for null plots and bootstrapped plots. The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots. The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$. (D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model. ', fig.height = 8}
p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle(glue::glue("(A) Residual plot (MVI = {format(my_vi$check_result$observed$vss - log(nrow(my_vi$get_fitted_and_resid())) + 10, digits = 3)}, strong)")) 
p2 <- readRDS(here::here("cached_data/boston_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(observed~hat(D)))) +
  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
pca_result <- my_vi$feature_pca()
pc1_per <- (attributes(pca_result)$sdev^2/256)[1]
pc2_per <- (attributes(pca_result)$sdev^2/256)[2]
p4 <- pca_result %>%
  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
  mutate(set = ifelse(set == "null", "Null", set)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = 7) +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  xlab(glue::glue("PC1 ({scales::percent(pc1_per)})")) +
  ylab(glue::glue("PC2 ({scales::percent(pc2_per)})")) +
  labs(col = "") +
  theme(legend.position = "bottom") +
  ggtitle("(D) PCA for extracted features") +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))

@@

### DatasauRus

@@

The "dino" dataset within the `datasauRus` R package exemplifies this scenario. With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur", as displayed in [@fig-dino-check]A. Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in @fig-dino-lineup. A visual test conducted by humans would undoubtedly reject $H_0$.

@@

```{r fig-dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the datasauRus fitted model evaluted on 200 null plots and 200 bootstrapped plots. (A) The residual plot exhibits a "dinosaur" shape. (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.  (C) The density plot of estimated distnace for null plots and bootstrapped plots. The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots. The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$. (D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model.', fig.height = 8}

p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle(glue::glue("(A) Residual plot (MVI = {format(my_vi$check_result$observed$vss - log(nrow(my_vi$get_fitted_and_resid())) + 10, digits = 3)}, strong)")) 
p2 <- readRDS(here::here("cached_data/dino_attention_map.rds")) + ggtitle("(B) Attention map")
p3 <- my_vi$summary_plot() +
  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("RESET test p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}\n Breusch-Pagan test p-value = {format(lmtest::bptest(mod)$p.value, digits = 3)} \n Shapiro-Wilk test p-value = {format(shapiro.test(mod$residuals)$p.value, digits = 3)}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(observed~hat(D)))) +
  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
pca_result <- my_vi$feature_pca()
pc1_per <- (attributes(pca_result)$sdev^2/256)[1]
pc2_per <- (attributes(pca_result)$sdev^2/256)[2]
p4 <- pca_result %>%
  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
  mutate(set = ifelse(set == "null", "Null", set)) %>%
  ggplot() +
  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = -10) +
  theme_light() +
  scale_color_brewer(palette = "Dark2") +
  xlab(glue::glue("PC1 ({scales::percent(pc1_per)})")) +
  ylab(glue::glue("PC2 ({scales::percent(pc2_per)})")) +
  labs(col = "", shape = "") +
  theme(legend.position = "bottom", legend.box="vertical") +
  ggtitle("(D) PCA for extracted features") +
  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))

@@

In this chapter, we have introduced a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference residual distribution assumed under correct model specification. This distance measure effectively captures the magnitude of model violations in misspecified models. We propose a computer vision model to estimate this distance, utilizing the residual plot of the fitted model as input. The resulting estimated distance serves as the foundation for constructing a single Model Violations Index (MVI), facilitating the quantification of various model violations.
